---
title: "Load data"
author: "Dave Edison Rojas Calderon"
date: "2023-04-06"
output: html_document
---
Functions
```{r}
filterTopN = function(RelAbTable, rel_abund_col, topN) {
  
  # Group by taxonomy and calculate the mean of relative abundance for each taxonomy
  taxa_mean = RelAbTable %>%
#    replace(is.na(.), 0) %>% 
    group_by(Taxonomy) %>%
    summarise(rel_abund_mean = mean(!!sym(rel_abund_col))) %>%
    ungroup()

  # Arrange by descending order of relative abundance and select the top N taxonomies
  top_taxa = taxa_mean %>%
    arrange(desc(rel_abund_mean)) %>%
    slice(1:topN)

  # Filter out the top N taxonomies and sum the relative abundance of the remaining taxa
  other_taxa = taxa_mean %>%
    anti_join(top_taxa, by = "Taxonomy") %>%
    summarise(Taxonomy = "Other") %>%
    mutate(rel_abund_mean = mean(RelAbTable[[rel_abund_col]][Taxonomy == "Other"]))

  # Combine the top N taxonomies and the "other" group
  final_df = bind_rows(top_taxa, other_taxa)

  # Join with original dataframe to get the sample indexes and time_points
  FilteredRelAbTable = RelAbTable %>%
    left_join(final_df, by = "Taxonomy") %>%
    mutate(Taxonomy = ifelse(is.na(rel_abund_mean), "Other", Taxonomy))

  return(FilteredRelAbTable)
}
```


Loading packages
```{r}
# Load library
library(qiime2R)
library(tidyverse)
library(gplots)
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(janitor)
library(patchwork)
library(gridExtra)
library(cowplot)
library(ggpubr)
library(forcats)
```

# Initialize of SVs, Taxonomy and Metadata
Open files
```{r}
# SVs
SVs = read_qza(file = "../data/salmon-5/table-atlanticsalmon-5.qza")

# Taxonomy assignations
taxonomy = read_qza(file = "../data/salmon-5/tax-class-atlanticsalmon-5.qza")

# Metadata
metadata = read_q2metadata("../data/salmon-5/metadata.tsv") %>% 
  mutate(phase2 = phase) %>% 
  mutate(WP2 = WP)

metadata = metadata %>% 
  mutate(SampleID = str_replace_all(SampleID, "-", ".")) %>% 
  mutate(phase2 = str_replace(phase2, "C0", "Observation")) %>% 
  mutate(phase2 = str_replace(phase2, "C1", "Intervention")) %>%
  mutate(WP2 = str_replace(WP2, "6A", "Wild")) %>% 
  mutate(WP2 = str_replace(WP2, "5A", "Farm")) %>% 
  rename(sample = SampleID)

meta = metadata %>% 
  select(sample, phase, phase2, WP, WP2, sample_site, country, timepoint, sample_dim, type_sample)

gc()
```
Count Table
```{r}
# SVs table
SVs = SVs$data %>% 
  data.frame() %>% 
  rownames_to_column(var = "Feature.ID")

# Taxonomy table
taxonomy = taxonomy$data %>% 
  parse_taxonomy()

# Add Unclassified if the row is empty
taxonomy[is.na(taxonomy)] = ""
taxonomy[taxonomy == "__"] = ""

for (i in 1:nrow(taxonomy)){
  if (taxonomy[i,7] != ""){
    taxonomy$Species[i] = paste(taxonomy$Genus[i], taxonomy$Species[i], sep = " ")
  } else if (taxonomy[i,2] == ""){
    kingdom = paste("Unclassified", taxonomy[i,1], sep = " ")
    taxonomy[i, 2:7] = kingdom
  } else if (taxonomy[i,3] == ""){
    phylum = paste("Unclassified", taxonomy[i,2], sep = " ")
    taxonomy[i, 3:7] = phylum
  } else if (taxonomy[i,4] == ""){
    class = paste("Unclassified", taxonomy[i,3], sep = " ")
    taxonomy[i, 4:7] = class
  } else if (taxonomy[i,5] == ""){
    order = paste("Unclassified", taxonomy[i,4], sep = " ")
    taxonomy[i, 5:7] = order
  } else if (taxonomy[i,6] == ""){
    family = paste("Unclassified", taxonomy[i,5], sep = " ")
    taxonomy[i, 6:7] = family
  } else if (taxonomy[i,7] == ""){
    taxonomy$Species[i] = paste("Unclassified ",taxonomy$Genus[i], sep = " ")
  }
}

# Add previous taxonomy level when "unclassified
# Iterate over each row of the dataframe
for (i in 1:nrow(taxonomy)) {
  # Iterate over each column of the row
  for (j in 1:ncol(taxonomy)) {
    # Check if the value is "uncultured"
    if (taxonomy[i, j] == "uncultured") {
      # Update the value with "uncultured" followed by the previous column value
      taxonomy[i, j] <- paste0("uncultured ", taxonomy[i, j - 1])
    }
  }
}

# Remove Eukaryota and Unnassigned
taxonomy_assigned = taxonomy %>% 
  filter(Kingdom != "d__Eukaryota") %>% 
  filter(Kingdom != "Unassigned")

# Save as a file
save(taxonomy_assigned,file="taxonomy_assigned.Rda")

gc()
```
Select phylum from taxonomy
```{r}
phylum = taxonomy_assigned %>% 
  rownames_to_column(var = "Feature.ID") %>% 
  select(Feature.ID, Phylum) %>% 
  right_join(SVs,
              by = "Feature.ID") %>% 
  select(!Feature.ID) %>% 
  drop_na() %>%
  group_by(Phylum) %>% 
  summarise_all(list(sum)) %>% 
  pivot_longer(names_to = "sample",
               values_to = "Count",
               cols = -Phylum) %>% 
  dplyr::select("sample", everything())
```
Join with metadata
```{r}
phylum = phylum %>% 
  inner_join(meta,
            by = join_by(sample)) %>% 
  rename(Taxonomy = Phylum)
```

# Filter and group in "Other"
Filter top 20 phyla across all samples
```{r}
phylum = phylum %>%
  mutate(Count = ifelse(is.na(Count), 0, Count)) %>% 
  group_by(sample) %>%
  mutate(relab_samples = Count / sum(Count) * 100) %>% 
  ungroup() %>% 
  mutate(relab_samples = ifelse(is.na(relab_samples), 0, relab_samples))

phylum_f = filterTopN(phylum, "relab_samples", 9)

phylum_f %>% 
  select(Taxonomy) %>% 
  unique()

phylum %>% 
  select(sample_dim) %>% 
  unique()
```
Order levels in sample type (diminutive - shorter name), time point, and taxonomy
```{r}
phylum_f$sample_dim = factor(phylum_f$sample_dim, levels = c("Fertilised eggs", "Yolk sack larvae", "Fry", "Skin", "Anterior intestine", "Mid intestine", "Posterior intestine", "Anterior digesta", "Mid digesta", "Filet", "Biofilm",  "Inlet water", "Tank water", "Outlet water", "River water", "Upstream water to farm", "Seawater inside pen", "Seawater outside pen", "Feed freshwater", "Feed sea cage"))

phylum_f$timepoint= factor(phylum_f$timepoint, levels = c("T0", "T1", "T2", "T3", "T4", "T5", "T6", "T7", "T8"))

phylum_f$Taxonomy = factor(phylum_f$Taxonomy, levels = c("Acidobacteriota", "Actinobacteriota", "Armatimonadota", "Bdellovibrionota", "Campilobacterota", "Chloroflexi", "Cyanobacteria", "Deinococcota", "Desulfobacterota", "Firmicutes", "Fusobacteriota", "Gemmatimonadota", "Halobacterota", "Nitrospirota", "Patescibacteria", "Planctomycetota", "Proteobacteria", "SAR324_clade(Marine_group_B)", "Verrucomicrobiota","WPS-2","Other"))
```

# Calculate relative abundance and mean of relative abundance
Group data
Mean relative abundance based on sample type
```{r}
# Group the data again by the same columns and calculate the mean
mean_data_type = phylum_f %>%
  group_by(country, WP, phase, type_sample, sample_dim, Taxonomy) %>%
  summarize(mean_relab = mean(relab_samples)) %>% 
  # Normalize mean abundances
  mutate(norm_mean_relab = mean_relab / sum(mean_relab) * 100)
```
Mean relative abundance based on time-point
```{r}
# Group the data again by the same columns and calculate the mean
mean_data_time = phylum_f %>%
  group_by(country, WP, phase, type_sample, timepoint, Taxonomy) %>%
  summarize(mean_relab = mean(relab_samples)) %>% 
  # Normalize mean abundances
  mutate(norm_mean_relab = mean_relab / sum(mean_relab) * 100)
```
# PLOT
Colors
```{r}
my_colors = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd",
                 "#8c564b", "#e377c2", "#7f7f7f", "#bcbd22", "#17becf",
                 "#ff9896", "#98df8a", "#f7b6d2", "#c5b0d5", "#c7c7c7",
                 "#dbdb8d", "#9edae5", "#aec7e8", "#ffbb78", "#d95f02",
                 "#fb8072", "#80b1d3", "#fdb462", "#b3de69", "#fccde5",
                 "#d9d9d9", "#bc80bd", "#ccebc5", "#ffed6f", "#1b9e77",
                 "#fdd0a2")
```
## IRELAND FISH
```{r}
phy_ire_5A_time = mean_data_time %>%
  filter(country == "Ireland") %>% 
  filter(WP == "5A") %>% 
  filter(type_sample == "Fish") %>%
  ggplot(aes(x = timepoint,
             y = norm_mean_relab,
             fill = Taxonomy)) + 
  geom_bar(stat = "identity",
           position = "stack") +
  theme_classic() +
  labs(y = "",
       x = "",
       fill = "Taxonomy") +
  scale_fill_manual(values = my_colors) +
  theme(axis.text.x = element_text(angle=0, vjust=1, hjust=1),
        legend.key.size = unit(1, 'cm'),
        legend.position = "none") +
  theme(text = element_text(size = 40))

phy_ire_6A_time = mean_data_time %>%
  filter(country == "Ireland") %>% 
  filter(WP == "6A") %>% 
  filter(type_sample == "Fish") %>%
  ggplot(aes(x = timepoint,
             y = norm_mean_relab,
             fill = Taxonomy)) + 
  geom_bar(stat = "identity",
           position = "stack") +
  theme_classic() +
  labs(y = "",
       x = "Time point",
       fill = "Taxonomy") +
  scale_fill_manual(values = my_colors) +
  theme(axis.text.x = element_text(angle=0, vjust=1, hjust=1),
        legend.key.size = unit(2, 'cm'),
        legend.position = "right") +
  theme(text = element_text(size = 40))

phy_ire_5A_samp = mean_data_type %>%
  filter(country == "Ireland") %>% 
  filter(WP == "5A") %>% 
  filter(type_sample == "Fish") %>% 
  ggplot(aes(x = sample_dim,
             y = norm_mean_relab,
             fill = Taxonomy)) + 
  geom_bar(stat = "identity",
           position = "stack") +
  theme_classic() +
  labs(y = "Relative Abundance (%)",
       x = "",
       title = "A") +
  scale_fill_manual(values = my_colors) +
  theme(axis.text.x = element_text(angle=90, vjust=1, hjust=1),
        legend.key.size = unit(2, 'cm'),
        legend.position = "none") +
  theme(text = element_text(size = 40))

phy_ire_6A_samp = mean_data_type %>%
  filter(country == "Ireland") %>% 
  filter(WP == "6A") %>% 
  filter(type_sample == "Fish") %>% 
  ggplot(aes(x = sample_dim,
             y = norm_mean_relab,
             fill = Taxonomy)) + 
  geom_bar(stat = "identity",
           position = "stack") +
  theme_classic() +
  labs(y = "Relative Abundance (%)",
       x = "Sample type",
       title = "C") +
  scale_fill_manual(values = my_colors) +
  theme(axis.text.x = element_text(angle=90, vjust=1, hjust=1),
        legend.key.size = unit(2, 'cm'),
        legend.position = "none") +
  theme(text = element_text(size = 40))
```

## IRELAND ENVIRONMENTAL
```{r}
phy_ire_5A_time_env = mean_data_time %>%
  filter(country == "Ireland") %>% 
  filter(WP == "5A") %>% 
  filter(type_sample == "Environmental") %>%
  ggplot(aes(x = timepoint,
             y = norm_mean_relab,
             fill = Taxonomy)) + 
  geom_bar(stat = "identity",
           position = "stack") +
  theme_classic() +
  labs(y = "",
       x = "",
       title = "") +
  scale_fill_manual(values = my_colors) +
  theme(axis.text.x = element_text(angle=0, vjust=1, hjust=1),
        legend.key.size = unit(2, 'cm'),
        legend.position = "none") +
  theme(text = element_text(size = 40))

phy_ire_5A_samp_env = mean_data_type %>%
  filter(country == "Ireland") %>% 
  filter(WP == "5A") %>% 
  filter(type_sample == "Environmental") %>% 
  ggplot(aes(x = sample_dim,
             y = norm_mean_relab,
             fill = Taxonomy)) + 
  geom_bar(stat = "identity",
           position = "stack") +
  theme_classic() +
  labs(y = "Relative Abundance (%)",
       x = "",
       title = "B") +
  scale_fill_manual(values = my_colors) +
  theme(axis.text.x = element_text(angle=90, vjust=1, hjust=1),
        legend.key.size = unit(2, 'cm'),
        legend.position = "none") +
  theme(text = element_text(size = 40))
```
Save plots
```{r}
phy_ire_5a = (phy_ire_5A_samp + phy_ire_5A_time) / (phy_ire_5A_samp_env + phy_ire_5A_time_env) / (phy_ire_6A_samp + phy_ire_6A_time)  + plot_layout(guides = "collect")

ggsave("../plots_no_unnassigned/barplots_mean_phylum/phy_ire_all.pdf", width = 25, height = 35)
ggsave("../plots_no_unnassigned/barplots_mean_phylum/phy_ire_all.png", width = 25, height = 35)
```
## NORWAY 5A
```{r}
phy_nor_c0_time = mean_data_time %>%
  filter(country == "Norway") %>% 
  filter(WP == "5A") %>% 
  filter(phase == "C0") %>% 
  filter(type_sample == "Fish") %>%
  ggplot(aes(x = timepoint,
             y = norm_mean_relab,
             fill = Taxonomy)) + 
  geom_bar(stat = "identity",
           position = "stack") +
  theme_classic() +
  labs(y = "",
       x = "") +
  scale_fill_manual(values = my_colors) +
  theme(axis.text.x = element_text(angle=0, vjust=1, hjust=1),
        legend.key.size = unit(1, 'cm'),
        legend.position = "none") +
  theme(text = element_text(size = 40))

phy_nor_c1_time = mean_data_time %>%
  filter(country == "Norway") %>% 
  filter(WP == "5A") %>% 
  filter(phase == "C1") %>% 
  filter(type_sample == "Fish") %>%
  ggplot(aes(x = timepoint,
             y = norm_mean_relab,
             fill = Taxonomy)) + 
  geom_bar(stat = "identity",
           position = "stack") +
  theme_classic() +
  labs(y = "",
       x = "") +
  scale_fill_manual(values = my_colors) +
  theme(axis.text.x = element_text(angle=0, vjust=1, hjust=1),
        legend.key.size = unit(2, 'cm'),
        legend.position = "right") +
  theme(text = element_text(size = 40))

phy_nor_c0_samp = mean_data_type %>%
  filter(country == "Norway") %>% 
  filter(WP == "5A") %>%
  filter(phase == "C0") %>% 
  filter(type_sample == "Fish") %>% 
  ggplot(aes(x = sample_dim,
             y = norm_mean_relab,
             fill = Taxonomy)) + 
  geom_bar(stat = "identity",
           position = "stack") +
  theme_classic() +
  labs(y = "Relative Abundance (%)",
       x = "",
       title = "A") +
  scale_fill_manual(values = my_colors) +
  theme(axis.text.x = element_text(angle=90, vjust=1, hjust=1),
        legend.key.size = unit(2, 'cm'),
        legend.position = "none") +
  theme(text = element_text(size = 40))

phy_nor_c1_samp = mean_data_type %>%
  filter(country == "Norway") %>% 
  filter(WP == "5A") %>% 
  filter(phase == "C1") %>% 
  filter(type_sample == "Fish") %>% 
  ggplot(aes(x = sample_dim,
             y = norm_mean_relab,
             fill = Taxonomy)) + 
  geom_bar(stat = "identity",
           position = "stack") +
  theme_classic() +
  labs(y = "Relative Abundance (%)",
       x = "",
       title = "B") +
  scale_fill_manual(values = my_colors) +
  theme(axis.text.x = element_text(angle=90, vjust=1, hjust=1),
        legend.key.size = unit(2, 'cm'),
        legend.position = "none") +
  theme(text = element_text(size = 40))
```

## NORWAY 5A ENVIRONMENTAL
```{r}
phy_nor_c0_time_env = mean_data_time %>%
  filter(country == "Norway") %>% 
  filter(WP == "5A") %>% 
  filter(phase == "C0") %>% 
  filter(type_sample == "Environmental") %>%
  ggplot(aes(x = timepoint,
             y = norm_mean_relab,
             fill = Taxonomy)) + 
  geom_bar(stat = "identity",
           position = "stack") +
  theme_classic() +
  labs(y = "Relative Abundance (%)",
       x = "Time point",
       title = "") +
  scale_fill_manual(values = my_colors) +
  theme(axis.text.x = element_text(angle=0, vjust=1, hjust=1),
        legend.key.size = unit(2, 'cm'),
        legend.position = "none") +
  theme(text = element_text(size = 40))

phy_nor_c0_samp_env = mean_data_type %>%
  filter(country == "Norway") %>% 
  filter(WP == "5A") %>% 
  filter(phase == "C0") %>% 
  filter(type_sample == "Environmental") %>% 
  ggplot(aes(x = sample_dim,
             y = norm_mean_relab,
             fill = Taxonomy)) + 
  geom_bar(stat = "identity",
           position = "stack") +
  theme_classic() +
  labs(y = "Relative Abundance (%)",
       x = "Sample type",
       title = "C") +
  scale_fill_manual(values = my_colors) +
  theme(axis.text.x = element_text(angle=90, vjust=1, hjust=1),
        legend.key.size = unit(2, 'cm'),
        legend.position = "none") +
  theme(text = element_text(size = 40))
```
Save plots
```{r}
phy_nor_5a = (phy_nor_c0_samp + phy_nor_c0_time) / (phy_nor_c1_samp + phy_nor_c1_time) / (phy_nor_c0_samp_env + phy_nor_c0_time_env) + plot_layout(guides = "collect")

ggsave("../plots_no_unnassigned/barplots_mean_phylum/phy_nor_all.pdf", width = 25, height = 35)
ggsave("../plots_no_unnassigned/barplots_mean_phylum/phy_nor_all.png", width = 25, height = 35)
```
# NORWAY 6A
```{r}
phy_nor_6a_time = mean_data_time %>%
  filter(country == "Norway") %>% 
  filter(WP == "6A") %>% 
  filter(type_sample == "Fish") %>%
  ggplot(aes(x = timepoint,
             y = norm_mean_relab,
             fill = Taxonomy)) + 
  geom_bar(stat = "identity",
           position = "stack") +
  theme_classic() +
  labs(y = "",
       x = "",
       title = "") +
  scale_fill_manual(values = my_colors) +
  theme(axis.text.x = element_text(angle=0, vjust=1, hjust=1),
        legend.key.size = unit(2, 'cm'),
        legend.position = "right") +
  theme(text = element_text(size = 40))

phy_nor_6a_time_env = mean_data_time %>%
  filter(country == "Norway") %>% 
  filter(WP == "6A") %>% 
  filter(type_sample == "Environmental") %>%
  ggplot(aes(x = timepoint,
             y = norm_mean_relab,
             fill = Taxonomy)) + 
  geom_bar(stat = "identity",
           position = "stack") +
  theme_classic() +
  labs(y = "",
       x = "Time point",
       title = "") +
  scale_fill_manual(values = my_colors) +
  theme(axis.text.x = element_text(angle=0, vjust=1, hjust=1),
        legend.key.size = unit(2, 'cm'),
        legend.position = "right") +
  theme(text = element_text(size = 40))

phy_nor_6a_samp = mean_data_type %>%
  filter(country == "Norway") %>% 
  filter(WP == "6A") %>% 
  filter(type_sample == "Fish") %>%
  ggplot(aes(x = sample_dim,
             y = norm_mean_relab,
             fill = Taxonomy)) + 
  geom_bar(stat = "identity",
           position = "stack") +
  theme_classic() +
  labs(y = "Relative Abundance (%)",
       x = "",
       title = "A") +
  scale_fill_manual(values = my_colors) +
  theme(axis.text.x = element_text(angle=90, vjust=1, hjust=1),
        legend.key.size = unit(2, 'cm'),
        legend.position = "none") +
  theme(text = element_text(size = 40))

phy_nor_6a_samp_env = mean_data_type %>%
  filter(country == "Norway") %>% 
  filter(WP == "6A") %>% 
  filter(type_sample == "Environmental") %>%
  ggplot(aes(x = sample_dim,
             y = norm_mean_relab,
             fill = Taxonomy)) + 
  geom_bar(stat = "identity",
           position = "stack") +
  theme_classic() +
  labs(y = "Relative Abundance (%)",
       x = "Sample type",
       title = "B") +
  scale_fill_manual(values = my_colors) +
  theme(axis.text.x = element_text(angle=90, vjust=1, hjust=1),
        legend.key.size = unit(2, 'cm'),
        legend.position = "none") +
  theme(text = element_text(size = 40))
```
Save plots
```{r}
phy_nor_6a = (phy_nor_6a_samp + phy_nor_6a_time) / (phy_nor_6a_samp_env + phy_nor_6a_time_env) + plot_layout(guides = "collect")

ggsave("../plots_no_unnassigned/barplots_mean_phylum/phy_nor_6a_all.pdf", width = 30, height = 30)
ggsave("../plots_no_unnassigned/barplots_mean_phylum/phy_nor_6a_all.png", width = 30, height = 30)
```
