---
title: "Beta dispersion. anova and HSD"
author: "Dave Edison Rojas Calderon"
date: "2023-06-30"
output: html_document
---

Load packages
```{r}
library(tidyverse)
library(stringr)
library(tidyr)
library(fs)
library(dplyr)
library(ggplot2)
library(Compositional)
library(broom)  # devtools::install_github("tidymodels/broom")
library(cowplot)
library(janitor)
library(qiime2R)
library(factoextra)
library(readxl)
library(vegan)
library(patchwork)
```

# FUNCTIONS
# Turkey's Honest Significant Differences function
```{r}
TurkeyHSD_test = function(betadisper_output){
  (mod.HSD <- TukeyHSD(betadisper_output))

  comparisons = rownames(mod.HSD$group)

  mod.HSD.df = mod.HSD$group %>% 
    as_tibble()

  row.names(mod.HSD.df) = comparisons

  # Add significance column
  mod.HSD.df = mod.HSD.df %>% 
    rownames_to_column(var = "comparisons") %>% 
    mutate(significance = "")

  # Iterate over each row in the tibble
  for (i in 1:nrow(mod.HSD.df)) {
    if (mod.HSD.df$`p adj`[i] < 0.05) {
      mod.HSD.df$significance[i] = "*"
    } else if (mod.HSD.df$`p adj`[i] < 0.01) {
      mod.HSD.df$significance[i] = "**"
    } else if (mod.HSD.df$`p adj`[i] < 0.001) {
      mod.HSD.df$significance[i] = "***"
    }
  }
  
  #mod.HSD.df = mod.HSD.df %>% 
    #filter(`p adj` <= 0.05)
  
  return(mod.HSD.df)
}
```

Load data
```{r}
# SVs
SVs = read_qza(file = "../data/salmon-5/table-atlanticsalmon-5.qza")

# Taxonomy assignations
taxonomy = read_qza(file = "../data/salmon-5/tax-class-atlanticsalmon-5.qza")

# Metadata
metadata = read_q2metadata("../data/salmon-5/metadata.tsv") %>% 
  mutate(phase2 = phase) %>% 
  mutate(WP2 = WP)

metadata = metadata %>% 
  mutate(SampleID = str_replace_all(SampleID, "-", ".")) %>% 
  mutate(phase2 = str_replace(phase2, "C0", "Observation")) %>% 
  mutate(phase2 = str_replace(phase2, "C1", "Intervention")) %>%
  mutate(WP2 = str_replace(WP2, "6A", "Wild")) %>% 
  mutate(WP2 = str_replace(WP2, "5A", "Farm")) %>% 
  rename(sample = SampleID)

meta = metadata %>% 
  select(sample, phase, phase2, WP, WP2, sample_group, country, timepoint, sample_dim, type_sample)
```
Clean and transform into a table
```{r}
# SVs table
SVs = SVs$data %>% 
  data.frame() %>% 
  rownames_to_column(var = "Feature.ID")

# Taxonomy table
taxonomy = taxonomy$data %>% 
  parse_taxonomy()

# Add Unclassified if the row is empty
taxonomy[is.na(taxonomy)] = ""
taxonomy[taxonomy == "__"] = ""

for (i in 1:nrow(taxonomy)){
  if (taxonomy[i,7] != ""){
    taxonomy$Species[i] <- paste(taxonomy$Genus[i], taxonomy$Species[i], sep = " ")
  } else if (taxonomy[i,2] == ""){
    kingdom <- paste("Unclassified", taxonomy[i,1], sep = " ")
    taxonomy[i, 2:7] <- kingdom
  } else if (taxonomy[i,3] == ""){
    phylum <- paste("Unclassified", taxonomy[i,2], sep = " ")
    taxonomy[i, 3:7] <- phylum
  } else if (taxonomy[i,4] == ""){
    class <- paste("Unclassified", taxonomy[i,3], sep = " ")
    taxonomy[i, 4:7] <- class
  } else if (taxonomy[i,5] == ""){
    order <- paste("Unclassified", taxonomy[i,4], sep = " ")
    taxonomy[i, 5:7] <- order
  } else if (taxonomy[i,6] == ""){
    family <- paste("Unclassified", taxonomy[i,5], sep = " ")
    taxonomy[i, 6:7] <- family
  } else if (taxonomy[i,7] == ""){
    taxonomy$Species[i] <- paste("Unclassified ",taxonomy$Genus[i], sep = " ")
  }
}

# Add previous taxonomy level when "unclassified
# Iterate over each row of the dataframe
for (i in 1:nrow(taxonomy)) {
  # Iterate over each column of the row
  for (j in 1:ncol(taxonomy)) {
    # Check if the value is "uncultured"
    if (taxonomy[i, j] == "uncultured") {
      # Update the value with "uncultured" followed by the previous column value
      taxonomy[i, j] <- paste0("uncultured ", taxonomy[i, j - 1])
    }
  }
}

# Remove Eukaryota
taxonomy = taxonomy %>% 
  filter(Kingdom != "d__Eukaryota")

gc()
```
Keep Genus level and join SVs table, zero replacement
```{r}
count = taxonomy %>% 
  rownames_to_column(var = "Feature.ID") %>% 
  select(Feature.ID, Genus) %>% 
  right_join(SVs,
              by = "Feature.ID") %>% 
  select(!Feature.ID) %>% 
  drop_na() %>%
  group_by(Genus) %>% 
  summarise_all(list(sum))

gc()
```
Transpose the table
```{r}
#  Save row names
rownames = colnames(count)

# Transpose
count = t(count) %>% 
  as_tibble()

# First row to column names
count = count %>% 
  row_to_names(row_number = 1)

# Add row names to transpose data
row.names(count) = rownames[-1]

# Row to column names
count = count %>% 
  rownames_to_column(var = "sample")
```
## Relative abundance
Data wide to long
```{r}
rel_ab = count %>% 
  pivot_longer(cols = !sample,
               names_to = "Genus",
               values_to = "count")
```
Calculate relative abundance
```{r}
rel_ab$count = as.numeric(as.character(rel_ab$count))

rel_ab = rel_ab %>% 
  group_by(sample) %>% 
  mutate(rel_ab = count / sum(count)) %>% 
  ungroup()
```
Pivot longer to wide
```{r}
rel_ab = rel_ab %>% 
  select(!count) %>% 
  pivot_wider(names_from = Genus,
              values_from = rel_ab) %>% 
  drop_na()
```

Join metadata to count table (General table)
```{r}
relab_meta = rel_ab %>% 
  left_join(meta,
            by = "sample") %>% 
  drop_na()

countries = relab_meta$country
sample_group = relab_meta$sample_group
```

Sub-tables for beta dispersion comparing
```{r}
# Ireland
relab_ire = relab_meta %>% 
  filter(country == "Ireland")
### Only relative abundance
only_relab_ire = relab_ire %>% 
  select(-c(sample, phase, phase2, WP, WP2, sample_group, country, timepoint, sample_dim, type_sample)) %>% 
  as.matrix()
### save grouping
wp_ire = relab_ire$WP

## WP5
relab_ire_wp5 = relab_ire %>% 
  filter(WP == "5A")
### Only relative abundance
only_relab_ire_wp5 = relab_ire_wp5 %>% 
  select(-c(sample, phase, phase2, WP, WP2, sample_group, country, timepoint, sample_dim, type_sample)) %>% 
  as.matrix()
### save grouping
samp_ire_wp5 = relab_ire_wp5$sample_dim
time_ire_wp5 = relab_ire_wp5$timepoint
group_ire_wp5 = relab_ire_wp5$sample_group

## WP6
relab_ire_wp6 = relab_ire %>% 
  filter(WP == "6A")
### Only relative abundance
only_relab_ire_wp6 = relab_ire_wp6 %>% 
  select(-c(sample, phase, phase2, WP, WP2, sample_group, country, timepoint, sample_dim, type_sample)) %>% 
  as.matrix()
### save grouping
samp_ire_wp6 = relab_ire_wp6$sample_dim
time_ire_wp6 = relab_ire_wp6$timepoint
group_ire_wp6 = relab_ire_wp6$sample_group

# Norway
relab_nor = relab_meta %>% 
  filter(country == "Norway")
### Only relative abundance
only_relab_nor = relab_nor %>% 
  select(-c(sample, phase, phase2, WP, WP2, sample_group, country, timepoint, sample_dim, type_sample)) %>% 
  as.matrix()
### save grouping
wp_nor = relab_nor$WP
phases_nor = relab_nor$phase

## WP5
relab_nor_wp5 = relab_nor %>% 
  filter(WP == "5A")
### Only relative abundance
only_relab_nor_wp5 = relab_nor_wp5 %>% 
  select(-c(sample, phase, phase2, WP, WP2, sample_group, country, timepoint, sample_dim, type_sample)) %>% 
  as.matrix()
### save grouping
samp_nor_wp5 = relab_nor_wp5$sample_dim
time_nor_wp5 = relab_nor_wp5$timepoint
group_nor_wp5 = relab_nor_wp5$sample_group

## WP6
relab_nor_wp6 = relab_nor %>% 
  filter(WP == "6A")
### Only relative abundance
only_relab_nor_wp6 = relab_nor_wp6 %>% 
  select(-c(sample, phase, phase2, WP, WP2, sample_group, country, timepoint, sample_dim, type_sample)) %>% 
  as.matrix()
### save grouping
samp_nor_wp6 = relab_nor_wp6$sample_dim
time_nor_wp6 = relab_nor_wp6$timepoint
group_nor_wp6 = relab_nor_wp6$sample_group
```

# Calculate Bray-Curtis dissimilarity
General
```{r}
only_relab = rel_ab %>% 
  select(!sample) %>% 
  as.matrix()

mode(only_relab) = "numeric"

set.seed(123)

dist = vegdist(only_relab, method = "bray")
```
Calculate for subgroups
```{r}
# Transform to numeric
mode(only_relab_ire) = "numeric"
mode(only_relab_ire_wp5) = "numeric"
mode(only_relab_ire_wp6) = "numeric"
mode(only_relab_nor) = "numeric"
mode(only_relab_nor_wp5) = "numeric"
mode(only_relab_nor_wp6) = "numeric"

# Set sed
set.seed(123)

# Calculate Bray-Curtis dissimilarity
dist_ire = vegdist(only_relab_ire, method = "bray")
dist_ire_wp5 = vegdist(only_relab_ire_wp5, method = "bray")
dist_ire_wp6 = vegdist(only_relab_ire_wp6, method = "bray")
dist_nor = vegdist(only_relab_nor, method = "bray")
dist_nor_wp5 = vegdist(only_relab_nor_wp5, method = "bray")
dist_nor_wp6 = vegdist(only_relab_nor_wp6, method = "bray")
```

# Calculate Beta dispersion
General
```{r}
mod = betadisper(dist, sample_group)
mod_countries = betadisper(dist, countries)
```
Calculate for subgroups
```{r}
# Ireland
## To compare WPs
mod_ire_wps = betadisper(dist_ire, wp_ire)

## WP5
### Sample type
mod_ire_wp5_samp = betadisper(dist_ire_wp5, samp_ire_wp5)
### Time points
mod_ire_wp5_time = betadisper(dist_ire_wp5, time_ire_wp5)
### Sample Group
mod_ire_wp5_group = betadisper(dist_ire_wp5, group_ire_wp5)

## WP6
### Sample type
mod_ire_wp6_samp = betadisper(dist_ire_wp6, samp_ire_wp6)
### Time points
mod_ire_wp6_time = betadisper(dist_ire_wp6, time_ire_wp6)
### Sample Group
mod_ire_wp6_group = betadisper(dist_ire_wp6, group_ire_wp6)


# Norway
## To compare WPs
mod_nor_wps = betadisper(dist_nor, wp_nor)
## To compare Phases
mod_nor_phases = betadisper(dist_nor, phases_nor)

## WP5
### Sample type
mod_nor_wp5_samp = betadisper(dist_nor_wp5, samp_nor_wp5)
### Time points
mod_nor_wp5_time = betadisper(dist_nor_wp5, time_nor_wp5)
### Sample Group
mod_nor_wp5_group = betadisper(dist_nor_wp5, group_nor_wp5)

## WP6
### Sample type
mod_nor_wp6_samp = betadisper(dist_nor_wp6, samp_nor_wp6)
### Time points
mod_nor_wp6_time = betadisper(dist_nor_wp6, time_nor_wp6)
### Sample Group
mod_nor_wp6_group = betadisper(dist_nor_wp6, group_nor_wp6)
```
Boxplots of beta dispersion
General
```{r}
## GENERAL
# Save plot
png(file="../plots/boxplot_beta disper/ire_nor.png", width=400, height=300, res = 100)

# Countries
boxplot(mod_countries, xlab = "Country")

dev.off()
```
Ireland
```{r}
## IRELAND
# Save plot
png(file="../plots/boxplot_beta disper/ire.png", width=700, height=700, res = 100)

# Combine plots
layout(matrix(c(1,1,2,3,4,5), 3, 2, byrow = TRUE))
# Ireland WPs
boxplot(mod_ire_wps, xlab = "Work package")

# WP5 sample grouped and time points
boxplot(mod_ire_wp5_group, xlab = "Sample type")
boxplot(mod_ire_wp5_time, xlab = "Time point")

# WP6 sample grouped and time points
boxplot(mod_ire_wp6_group, xlab = "Sample type")
boxplot(mod_ire_wp6_time, xlab = "Time point")

dev.off()

## NORWAY
# Save plot
png(file="../plots/boxplot_beta disper/nor.png", width=700, height=700, res = 100)

# Combine plots
layout(matrix(c(1,2,3,4,5,6), 3, 2, byrow = TRUE))

# Norway WPs and phases
boxplot(mod_nor_wps, xlab = "Work package")
boxplot(mod_nor_phases, xlab = "Phases")

# WP5 sample grouped and time points
boxplot(mod_nor_wp5_group, xlab = "Sample type")
boxplot(mod_nor_wp5_time, xlab = "Time point")

# WP6 sample grouped and time points
boxplot(mod_nor_wp6_group, xlab = "Sample type")
boxplot(mod_nor_wp6_time, xlab = "Time point")

dev.off()
```

# Perform test
General
```{r}
anova(mod)
anova(mod_countries)
```
For subgroups from Ireland
```{r}
print("Ireland comparing WPs")
anova(mod_ire_wps)
print("Ireland WP5 comparing sample type")
anova(mod_ire_wp5_samp)
print("Ireland WP5 comparing time-points")
anova(mod_ire_wp5_time)
print("Ireland WP5 comparing sample group")
anova(mod_ire_wp5_group)
print("Ireland WP6 comparing sample type")
anova(mod_ire_wp6_samp)
print("Ireland WP6 comparing time-point")
anova(mod_ire_wp6_time)
print("Ireland WP6 comparing sample group")
anova(mod_ire_wp6_group)
```

For subgroups from Norway
```{r}
print("Norway comparing WPs")
anova(mod_nor_wps)
print("Norway comparing phases")
anova(mod_nor_phases)
print("Norway WP5 comparing sample type")
anova(mod_nor_wp5_samp)
print("Norway WP5 comparing time-points")
anova(mod_nor_wp5_time)
print("Norway WP5 comparing sample group")
anova(mod_nor_wp5_group)
print("Norway WP6 comparing sample type")
anova(mod_nor_wp6_samp)
print("Norway WP6 comparing time-points")
anova(mod_nor_wp6_time)
print("Norway WP5 comparing sample group")
anova(mod_nor_wp6_group)
```


# Permutation test for F
```{r}
#permutest(mod, pairwise = TRUE, permutations = 99
```

# TurkeyHSD test
General
```{r}
# Sample groups
turkey_samp = TurkeyHSD_test(mod)
write.csv(turkey_samp, "../turkey_samp.csv", row.names=FALSE)
```
For subgroups in Ireland
```{r}
TurkeyHSD_test(mod_ire_wp5_samp)

TurkeyHSD_test(mod_ire_wp5_group)

TurkeyHSD_test(mod_ire_wp6_samp)

TurkeyHSD_test(mod_ire_wp6_time)

TurkeyHSD_test(mod_ire_wp6_group)
```
For subgroups in Norway
```{r}
TurkeyHSD_test(mod_nor_wp5_samp)

TurkeyHSD_test(mod_nor_wp5_time)

TurkeyHSD_test(mod_nor_wp5_group)

TurkeyHSD_test(mod_nor_wp6_samp)

TurkeyHSD_test(mod_nor_wp6_group)
```



