---
title: "NMDS_based_on_abundance"
author: "Dave Edison Rojas Calderon"
date: "2023-04-16"
output: html_document
---

Load libraries
```{r}
library(tidyverse)
library(stringr)
library(tidyr)
library(fs)
library(dplyr)
library(ggplot2)
library(Compositional)
library(broom)  # devtools::install_github("tidymodels/broom")
library(cowplot)
library(janitor)
library(qiime2R)
library(factoextra)
library(readxl)
library(vegan)
library(patchwork)
```

# DATA WRANGLING
Open files
```{r}
# SVs
SVs = read_qza(file = "../data/salmon-5/table-atlanticsalmon-5.qza")

# Taxonomy assignations
taxonomy = read_qza(file = "../data/salmon-5/tax-class-atlanticsalmon-5.qza")

# Metadata
metadata = read_q2metadata("../data/salmon-5/metadata.tsv") %>% 
  mutate(phase2 = phase) %>% 
  mutate(WP2 = WP)

metadata = metadata %>% 
  mutate(SampleID = str_replace_all(SampleID, "-", ".")) %>% 
  mutate(phase2 = str_replace(phase2, "C0", "Observation")) %>% 
  mutate(phase2 = str_replace(phase2, "C1", "Intervention")) %>%
  mutate(WP2 = str_replace(WP2, "6A", "Wild")) %>% 
  mutate(WP2 = str_replace(WP2, "5A", "Farm")) %>% 
  rename(sample = SampleID)

meta = metadata %>% 
  select(sample, phase, phase2, WP, WP2, sample_group, country, timepoint, sample_dim, type_sample)

save(meta, file = "meta.Rda")
```
Transform into tables
```{r}
# SVs table
SVs = SVs$data %>% 
  data.frame() %>% 
  rownames_to_column(var = "Feature.ID")

# Taxonomy table
taxonomy = taxonomy$data %>% 
  parse_taxonomy()

# Add Unclassified if the row is empty
taxonomy[is.na(taxonomy)] = ""
taxonomy[taxonomy == "__"] = ""

for (i in 1:nrow(taxonomy)){
  if (taxonomy[i,7] != ""){
    taxonomy$Species[i] <- paste(taxonomy$Genus[i], taxonomy$Species[i], sep = " ")
  } else if (taxonomy[i,2] == ""){
    kingdom <- paste("Unclassified", taxonomy[i,1], sep = " ")
    taxonomy[i, 2:7] <- kingdom
  } else if (taxonomy[i,3] == ""){
    phylum <- paste("Unclassified", taxonomy[i,2], sep = " ")
    taxonomy[i, 3:7] <- phylum
  } else if (taxonomy[i,4] == ""){
    class <- paste("Unclassified", taxonomy[i,3], sep = " ")
    taxonomy[i, 4:7] <- class
  } else if (taxonomy[i,5] == ""){
    order <- paste("Unclassified", taxonomy[i,4], sep = " ")
    taxonomy[i, 5:7] <- order
  } else if (taxonomy[i,6] == ""){
    family <- paste("Unclassified", taxonomy[i,5], sep = " ")
    taxonomy[i, 6:7] <- family
  } else if (taxonomy[i,7] == ""){
    taxonomy$Species[i] <- paste("Unclassified ",taxonomy$Genus[i], sep = " ")
  }
}

# Add previous taxonomy level when "unclassified
# Iterate over each row of the dataframe
for (i in 1:nrow(taxonomy)) {
  # Iterate over each column of the row
  for (j in 1:ncol(taxonomy)) {
    # Check if the value is "uncultured"
    if (taxonomy[i, j] == "uncultured") {
      # Update the value with "uncultured" followed by the previous column value
      taxonomy[i, j] <- paste0("uncultured ", taxonomy[i, j - 1])
    }
  }
}

# Remove Eukaryota
taxonomy = taxonomy %>% 
  filter(Kingdom != "d__Eukaryota")

gc()
```
Check the unique values in each taxonomy level to decide which one will be used for the NMDS
```{r}
# Phylum
taxonomy %>% 
  select(Phylum) %>% 
  unique()

# Class
taxonomy %>% 
  select(Class) %>% 
  unique()

# Order
taxonomy %>% 
  select(Order) %>% 
  unique()

# Genus
taxonomy %>% 
  select(Genus) %>% 
  unique()

# Species
taxonomy %>% 
  select(Species) %>% 
  unique()
```
Keep Genus level and join SVs table, zero replacement
```{r}
count_nmds = taxonomy %>% 
  rownames_to_column(var = "Feature.ID") %>% 
  select(Feature.ID, Genus) %>% 
  right_join(SVs,
              by = "Feature.ID") %>% 
  select(!Feature.ID) %>% 
  drop_na() %>%
  group_by(Genus) %>% 
  summarise_all(list(sum))

gc()
```
Transpose the table
```{r}
#  Save row names
rownames = colnames(count_nmds)

# Transpose
count_nmds = t(count_nmds) %>% 
  as_tibble()

# First row to column names
count_nmds = count_nmds %>% 
  row_to_names(row_number = 1)

# Add row names to transpose data
row.names(count_nmds) = rownames[-1]

# Row to column names
count_nmds = count_nmds %>% 
  rownames_to_column(var = "sample")
```
## Relative abundance
Data wide to long
```{r}
rel_ab = count_nmds %>% 
  pivot_longer(cols = !sample,
               names_to = "Genus",
               values_to = "count")
```
Calculate relative abundance
```{r}
rel_ab$count = as.numeric(as.character(rel_ab$count))

rel_ab = rel_ab %>% 
  group_by(sample) %>% 
  mutate(rel_ab = count / sum(count)) %>% 
  ungroup()
```
Pivot longer to wide
```{r}
rel_ab = rel_ab %>% 
  select(!count) %>% 
  pivot_wider(names_from = Genus,
              values_from = rel_ab) %>% 
  drop_na()
```

# NMDS (Non-metric Multi-dimensional Scaling)
Turn abundance data frame into a matrix
```{r}
m_com = rel_ab %>% 
  select(!sample) %>% 
  as.matrix()
```
Run metaMDS
```{r}
set.seed(123)
nmds = metaMDS(m_com, distance = "bray")
nmds
```
# First look
Stress plot
```{r}
stressplot(nmds)
```
Basic plot
```{r}
plot(nmds)
```
Extract NMDS scores (x and y coordinates) for sites from newer versions of vegan package
```{r}
data.scores = as.data.frame(scores(nmds)$sites)
```
Join to metadata
```{r}
data.scores$sample = rel_ab$sample

data.scores = data.scores %>% 
  inner_join(meta,
             by = "sample")

data.scores %>% 
  select(sample_group) %>% 
  unique()
```
Assign levels to sample type
```{r}
data.scores$sample_dim = factor(data.scores$sample_dim, levels = c("Fertilised eggs", "Yolk sack larvae", "Fry", "Skin", "Anterior intestine", "Mid intestine", "Posterior intestine", "Anterior digesta", "Mid digesta", "Filet", "Biofilm freshwater",  "Inlet water", "Tank water", "Outlet water", "Biofilm", "River water", "Biofilm sea cage", "Upstream water to farm", "Seawater inside pen", "Seawater outside pen", "Feed freshwater", "Feed sea cage"))

data.scores$sample_group = factor(data.scores$sample_group, levels = c("Early stages", "Skin", "Intestine", "Digesta", "Filet", "Biofilm",  "Fresh water", "Sea water", "River", "Feed"))
```
Assign levels to time point
```{r}
data.scores$timepoint= factor(data.scores$timepoint, levels = c("T0", "T1", "T2", "T3", "T4", "T5", "T6", "T7", "T8"))
```


# PLOT
Colors
```{r}
my_colors = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd",
             "#8c564b", "#e377c2", "#7f7f7f", "#bcbd22", "#17becf",
             "#ff9896", "#98df8a", "#f7b6d2", "#c5b0d5", "#c7c7c7",
             "#dbdb8d", "#9edae5", "#aec7e8")
```
## GENERAL FOR DISCUSSION
```{r}
# Rename sample group
data.scores = data.scores %>% 
  mutate(`Sample type` = sample_group)

nmds_general = data.scores %>% 
  ggplot(aes(x = NMDS1, 
             y = NMDS2, 
             color = `Sample type`)) +
  theme_bw() +
  scale_shape_manual(values = c(15, 16),
                     name="Source") +
  labs(title = "",
       x = "NMDS1",
       y = "NMDS2") +
  theme(text = element_text(size = 20))+
  stat_ellipse(geom = "polygon",
               aes(fill = `Sample type`), 
               alpha = 0.25)

ggsave("../plots/nmds_grouped/nmds_general.pdf", width = 12, height = 10)
ggsave("../plots/nmds_grouped/nmds_general.png", width = 12, height = 10) 
```

## Ireland WP5
```{r}
ire_wp5_bray_samp = data.scores %>%
  filter(country == "Ireland") %>%
  filter(WP == "5A") %>%
  filter(phase == "C0") %>% 
  ggplot(aes(x = NMDS1, 
             y = NMDS2, 
             color = sample_group,
             shape = type_sample)) +
  geom_point(alpha = 0.5,
             size = 5) +
  theme_bw() +
  scale_shape_manual(values = c(15, 16),
                     name="Source") +
  scale_size_continuous(name="Shannon Diversity") +
  scale_color_manual(values = my_colors, name = "Sample type") +
  labs(title = "A",
       x = "NMDS1",
       y = "NMDS2") +
  theme(text = element_text(size = 20))+
  stat_ellipse()

ire_wp5_bray_time = data.scores %>%
  filter(country == "Ireland") %>%
  filter(WP == "5A") %>%
  filter(phase == "C0") %>% 
  ggplot(aes(x = NMDS1, 
             y = NMDS2, 
             color = timepoint,
             shape = type_sample)) +
  geom_point(alpha = 0.5,
             size = 5) +
  theme_bw() +
  scale_shape_manual(values = c(15, 16),
                     name="Source") +
  scale_size_continuous(name="Shannon Diversity") +
  scale_color_manual(values = my_colors, name = "Time point") +
  labs(title = "A",
       x = "NMDS1",
       y = "NMDS2") +
  theme(text = element_text(size = 20))

ire_wp5_bray = ire_wp5_bray_samp / ire_wp5_bray_time

ggsave("../plots/nmds_grouped/ire_bray_wp5.pdf", width = 10, height = 12)
ggsave("../plots/nmds_grouped/ire_bray_wp5.png", width = 10, height = 12)    
```
## Ireland WP6
```{r}
ire_wp6_bray_samp = data.scores %>%
  filter(country == "Ireland") %>%
  filter(WP == "6A") %>%
  filter(phase == "C0") %>% 
  ggplot(aes(x = NMDS1, 
             y = NMDS2, 
             color = sample_group)) +
  geom_point(alpha=0.5,
             size = 5) +
  theme_bw() +
  scale_size_continuous(name="Shannon Diversity") +
  scale_color_manual(values = my_colors, name = "Sample type") +
  labs(title = "B",
       x = "NMDS1",
       y = "NMDS2") +
  theme(text = element_text(size = 20)) +
  stat_ellipse()

ire_wp6_bray_time = data.scores %>%
  filter(country == "Ireland") %>%
  filter(WP == "6A") %>%
  filter(phase == "C0") %>% 
  ggplot(aes(x = NMDS1, 
             y = NMDS2, 
             color = timepoint)) +
  geom_point(alpha=0.5,
             size = 5) +
  theme_bw() +
  scale_size_continuous(name="Shannon Diversity") +
  scale_color_manual(values = my_colors, name = "Time point") +
  labs(title = "B",
       x = "NMDS1",
       y = "NMDS2") +
  theme(text = element_text(size = 20))  

ire_wp6_bray = ire_wp6_bray_samp + ire_wp6_bray_time

ggsave("../plots/nmds_grouped/ire_bray_wp6.pdf", width = 10, height = 12)
ggsave("../plots/nmds_grouped/ire_bray_wp6.png", width = 10, height = 12)
```
Save
```{r}
# Sample groups
ire_samp = ire_wp5_bray_samp / ire_wp6_bray_samp

ggsave("../plots/nmds_grouped/ire_samp.pdf", width = 10, height = 12)
ggsave("../plots/nmds_grouped/ire_samp.png", width = 10, height = 12)

# Time points
ire_time = ire_wp5_bray_time / ire_wp6_bray_time

ggsave("../plots/nmds_grouped/ire_time.pdf", width = 10, height = 12)
ggsave("../plots/nmds_grouped/ire_time.png", width = 10, height = 12)
```


## Norway
### Phase c0 wp5
```{r}
nor_bray_c0_samp_wp5 = data.scores %>%
  filter(country == "Norway") %>%
  filter(phase == "C0") %>%
  filter(WP == "5A") %>% 
  ggplot(aes(x = NMDS1, 
             y = NMDS2, 
             color = sample_group,
             shape = type_sample)) +
  geom_point(alpha = 0.5,
             size = 5) +
  theme_bw() +
  scale_shape_manual(values = c(15, 16, 17, 18, 9, 10, 21, 22, 11, 0, 5, 6),
                     name="Source") +
  scale_size_continuous(name="Shannon Diversity") +
  scale_color_manual(values = my_colors, name = "Sample type") +
  labs(title = "A",
       x = "NMDS1",
       y = "NMDS2") +
  theme(text = element_text(size = 20)) +
  stat_ellipse()

nor_bray_c0_time_wp5 = data.scores %>%
  filter(country == "Norway") %>%
  filter(phase == "C0") %>%
  filter(WP == "5A") %>% 
  ggplot(aes(x = NMDS1, 
             y = NMDS2, 
             color = timepoint, 
             shape = type_sample,
             size = shannon_entropy)) +
  geom_point(alpha = 0.5,
             size = 5) +
  theme_bw() +
  scale_shape_manual(values = c(15, 16, 17, 18, 9, 10, 21, 22, 11, 0, 5, 6),
                     name="Source") +
  scale_size_continuous(name="Shannon Diversity") +
  scale_color_manual(values = my_colors, name = "Time point") +
  labs(title = "A",
       x = "NMDS1",
       y = "NMDS2") +
  theme(text = element_text(size = 20))  

nor_c0_pcoa = nor_bray_c0_samp_wp5 + nor_bray_c0_time_wp5

ggsave("../plots/nmds_grouped/nor_bray_c0_wp5.pdf", width = 10, height = 12)
ggsave("../plots/nmds_grouped/nor_bray_c0_wp5.png", width = 10, height = 12)
```

### Phase c0 wp6
```{r}
nor_bray_c0_samp_wp6 = data.scores %>%
  filter(country == "Norway") %>%
  filter(phase == "C0") %>%
  filter(WP == "6A") %>%
  ggplot(aes(x = NMDS1, 
             y = NMDS2, 
             color = sample_group,
             shape = type_sample)) +
  geom_point(alpha = 0.5,
             size = 5) +
  theme_bw() +
  scale_shape_manual(values = c(15, 16, 17, 18, 9, 10, 21, 22, 11, 0, 5, 6),
                     name="Source") +
  scale_size_continuous(name="Shannon Diversity") +
  scale_color_manual(values = my_colors, name = "Sample type") +
  labs(title = "C",
       x = "NMDS1",
       y = "NMDS2") +
  theme(text = element_text(size = 20)) +
  stat_ellipse()

nor_bray_c0_time_wp6 = data.scores %>%
  filter(country == "Norway") %>%
  filter(phase == "C0") %>%
  filter(WP == "6A") %>% 
  ggplot(aes(x = NMDS1, 
             y = NMDS2, 
             color = timepoint, 
             shape = type_sample)) +
  geom_point(alpha = 0.5,
             size = 5) +
  theme_bw() +
  scale_shape_manual(values = c(15, 16, 17, 18, 9, 10, 21, 22, 11, 0, 5, 6),
                     name="Source") +
  scale_color_manual(values = my_colors, name = "Time point") +
  labs(title = "C",
       x = "NMDS1",
       y = "NMDS2") +
  theme(text = element_text(size = 20))  

nor_c0_pcoa_wp6 = nor_bray_c0_samp_wp6 / nor_bray_c0_time_wp6

ggsave("../plots/nmds_grouped/nor_bray_c0_wp6.pdf", width = 10, height = 12)
ggsave("../plots/nmds_grouped/nor_bray_c0_wp6.png", width = 10, height = 12)
```
- Phase C1
```{r}
nor_bray_c1_samp_wp5 = data.scores %>%
  filter(country == "Norway") %>%
  filter(phase == "C1") %>%
  filter(WP == "5A") %>% 
  ggplot(aes(x = NMDS1, 
             y = NMDS2, 
             color = sample_group,
             shape = type_sample)) +
  geom_point(alpha = 0.5,
             size = 5) +
  theme_bw() +
  scale_shape_manual(values = c(15, 16, 17, 18, 9, 10, 21, 22, 11, 0, 5, 6),
                     name="Source") +
  scale_size_continuous(name="Shannon Diversity") +
  scale_color_manual(values = my_colors, name = "Sample type") +
  labs(title = "B",
       x = "NMDS1",
       y = "NMDS2") +
  theme(text = element_text(size = 20)) +
  stat_ellipse()

nor_bray_c1_time_wp5 = data.scores %>%
  filter(country == "Norway") %>%
  filter(phase == "C1") %>%
  filter(WP == "5A") %>% 
  ggplot(aes(x = NMDS1, 
             y = NMDS2, 
             color = timepoint, 
             shape = type_sample)) +
  geom_point(alpha = 0.5,
             size = 5) +
  theme_bw() +
  scale_shape_manual(values = c(15, 16, 17, 18, 9, 10, 21, 22, 11, 0, 5, 6),
                     name="Source") +
  scale_color_manual(values = my_colors, name = "Time point") +
  labs(title = "B",
       x = "NMDS1",
       y = "NMDS2") +
  theme(text = element_text(size = 20))  

nor_c1_pcoa = nor_bray_c1_samp_wp5 + nor_bray_c1_time_wp5

ggsave("../plots/nmds_grouped/nor_bray_c1_wp5.pdf", width = 10, height = 12)
ggsave("../plots/nmds_grouped/nor_bray_c1_wp5.png", width = 10, height = 12)
```
Save
```{r}
# Sample groups
nor_samp = nor_bray_c0_samp_wp5 / nor_bray_c1_samp_wp5 / nor_bray_c0_samp_wp6

ggsave("../plots/nmds_grouped/nor_samp.pdf", width = 10, height = 16)
ggsave("../plots/nmds_grouped/nor_samp.png", width = 10, height = 16)

# Time points
nor_time = nor_bray_c0_time_wp5 / nor_bray_c1_time_wp5 / nor_bray_c0_time_wp6

ggsave("../plots/nmds_grouped/nor_time.pdf", width = 10, height = 16)
ggsave("../plots/nmds_grouped/nor_time.png", width = 10, height = 16)
```

- Phase and WP
```{r}
nor_bray_phase = data.scores %>%
  filter(country == "Norway") %>%
  filter(WP == "5A") %>% 
  ggplot(aes(x = NMDS1, 
             y = NMDS2, 
             color = phase2)) +
  geom_point(alpha=0.5,
             size = 5) +
  theme_bw() +
  scale_color_discrete(name = "Phase") +
  theme(text = element_text(size = 20)) +
  labs(title = "A",
       x = "NMDS1",
       y = "NMDS2") +
  stat_ellipse()

nor_bray_wp = data.scores %>%
  filter(country == "Norway") %>%
  ggplot(aes(x = NMDS1, 
             y = NMDS2, 
             color = WP2)) +
  geom_point(alpha=0.5,
             size = 5) +
  theme_bw() +
  scale_color_discrete(name = "Work package") +
  theme(text = element_text(size = 20)) +
  labs(title = "B",
       x = "NMDS1",
       y = "NMDS2") +
  stat_ellipse()

nor_bray_wp_phases = nor_bray_phase / nor_bray_wp

ggsave("../plots/nmds_grouped/nor_bray_phases_wp.pdf", width = 10, height = 12)
ggsave("../plots/nmds_grouped/nor_bray_phases_wp.png", width = 10, height = 12)
```
- Countries and wp
```{r}
pcoa_country_wp = data.scores %>%
  ggplot(aes(x = NMDS1, 
             y = NMDS2, 
             color = country,
             shape = WP2)) +
  geom_point(alpha = 0.5,
             size = 5) +
  theme_bw() +
  scale_color_manual(values = c("brown3", "dodgerblue"), name="Country") +
   scale_shape_manual(values = c(15, 16, 17, 18, 9, 10, 21, 22, 11, 0, 5, 6),
                     name="Work Package") +
  labs(title = "Sample sites and WP",
       x = "NMDS1",
       y = "NMDS2") +
  theme(text = element_text(size = 20)) +
  stat_ellipse()

ggsave("../plots/nmds_grouped/nmds_country_wp.pdf", width = 10, height = 6)
ggsave("../plots/nmds_grouped/nmds_country_wp.png", width = 10, height = 6)
```

- countries and type of samples
```{r}
pcoa_country_wp = data.scores %>%
  ggplot(aes(x = NMDS1, 
             y = NMDS2, 
             color = country,
             shape = type_sample)) +
  geom_point(alpha = 0.5,
             size = 5) +
  theme_bw() +
  scale_color_manual(values = c("brown3", "dodgerblue"), name="Country") +
   scale_shape_manual(values = c(15, 16, 17, 18, 9, 10, 21, 22, 11, 0, 5, 6),
                     name="Type of sample") +
  labs(title = "Sample sites and type",
       x = "NMDS1",
       y = "NMDS2") +
  theme(text = element_text(size = 20)) +
  stat_ellipse()

ggsave("../plots/nmds_grouped/nmds_country_type.pdf", width = 10, height = 6)
ggsave("../plots/nmds_grouped/nmds_country_type.png", width = 10, height = 6)
```
